\textbf{ Interpretable Gaussian Processes for Stellar Light Curves }

### 2020-07-29 tex/ms.tex
second moment of rb dist
typos; finished mom2

\subsection{Distributions over $r$ and $\delta$}
%
We model the normalized spot radius $r$ as a random variable drawn from
a Beta distribution, whose PDF is given by
%
\begin{align}
    p \big(r \, \big| \, \alpha, \beta \big)
     & =
    \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
    r^{\alpha - 1}
    (1 - r)^{\beta - 1}
    \quad,
\end{align}
%
where $\Gamma$ is the Gamma function and $\alpha$ and $\beta$ are
hyperparameters characterizing the shape of the distribution.
%
Since we require the brghtness $b = 1 - \delta$ at the center of the spot to
be non-negative, it is convenient to parametrize the distribution of $b$
instead of $\delta$. We model $b$ as a random variable drawn from a log-normal
distribution:
%
\begin{align}
    p \big(b \, \big| \, \mu, \nu \big)
     & =
    \frac{1}{b\sqrt{2\pi\nu}}
    \exp\left[
        -\dfrac{\left(\ln b - \mu\right)^2}{2\nu}
        \right]
    \quad,
\end{align}
%
where $\mu$ and $\nu$ are the mean and variance of the log-normal, respectively.
The joint distribution over $r$ and $b$ is therefore separable,
by construction:
%
\begin{align}
    p \big( r, b \, \big| \, \alpha, \beta, \mu, \nu \big)
     & =
    p \big(r \, \big| \, \alpha, \beta \big)
    \,
    p \big(b \, \big| \, \mu, \nu \big)
    \quad.
\end{align}
%

\subsection{First moment}

@@

The first moment of the distribution of $y_{lm}$ over $r$ and $b$ is

@@

    \label{eq:Erb}

@@

        \mathrm{E} \big[ y_{lm} \big] &=

@@

@@

        \mathrm{E}_r \big[ y_{lm} \big]

@@

    \label{eq:Er}

@@

Expanding Equation~(\ref{eq:Er}), we obtain

@@

        \right)

@@

        l = m = 0    \\[2em]

@@

        \dfrac{(1 - b) c r \left( 2 + c r \right)r^{\alpha - 1}

@@

        l > 0, m = 0 \\[2em]

@@

        \qquad m \ne 0

@@

Thanks to the dependence of the spherical harmonic coefficients on only
powers of $r$ and $(1 + cr)$, the integrals above may be expressed in closed
form in terms of the hypergeometric function ${_2F_1}$:

@@

        1 -
        \dfrac{(1 - b) \alpha c}{2(\alpha + \beta)} F_l

@@

@@

        l = m = 0    \\[2em]
        -\dfrac{(1 - b) \alpha c}{\sqrt{2l + 1} (\alpha + \beta)}
        \left[
            F_l + \dfrac{(\alpha + 1)c}{2(\alpha + \beta + 1)} G_l
            \right]

@@

@@

        l > 0, m = 0 \\[2em]

@@

where we define
        F_l & \equiv {_2F_1}(l + 1, \alpha + 1; \alpha + \beta + 1; -c)
        \\[1em]
        G_l & \equiv {_2F_1}(l + 1, \alpha + 2; \alpha + \beta + 2; -c)
Note that for a given value of $\alpha$, $\beta$, and $c$, we need only compute
$F_0$, $F_1$, $G_0$, and $G_1$ directly, since the remaining terms may be
obtained recursively:
        F_l &=
        \dfrac{\alpha + \beta + 1 - l}{2l}F_{l - 2}
        - \dfrac{2\alpha + \beta + 2 - 3l}{2l}F_{l - 1}
        \\[1em]
        G_l &=
        \dfrac{\alpha + \beta + 2 - l}{2l}G_{l - 2}
        - \dfrac{2\alpha + \beta + 4 - 3l}{2l}G_{l - 1}
We may now evaluate the integral in Equation~(\ref{eq:Erb}):
    \mathrm{E} \big[ y_{lm} \big] & =
        \frac{1}{b\sqrt{2\pi\nu}}
        1 -
        \dfrac{(1 - b) \alpha c}{2(\alpha + \beta)} F_l
        \right)
        \qquad\qquad\times
        \exp\left[
            -\dfrac{\left(\ln b - \mu\right)^2}{2\nu}
            \right]
        \qquad\qquad
        l = m = 0
        -
        \frac{1}{b\sqrt{2\pi\nu}}
        \dfrac{(1 - b) \alpha c}{\sqrt{2l + 1} (\alpha + \beta)}
        \qquad\qquad\times
        \left[
            F_l + \dfrac{(\alpha + 1)c}{2(\alpha + \beta + 1)} G_l
            \right]
        \qquad\qquad\times
        \exp\left[
            -\dfrac{\left(\ln b - \mu\right)^2}{2\nu}
            \right]
        \qquad\qquad
        l > 0, m = 0 \\[2em]
        \qquad\qquad m \ne 0
Most of the terms in the integrands above are constants, so these
integrals may also be expressed in closed form:
    \mathrm{E} \big[ y_{lm} \big]
        1 -
        \dfrac{\gamma \alpha c}{2(\alpha + \beta)} F_l
        l = m = 0    \\[2em]
        -\dfrac{\gamma \alpha c}{\sqrt{2l + 1} (\alpha + \beta)}
        \left[
            F_l + \dfrac{(\alpha + 1)c}{2(\alpha + \beta + 1)} G_l
            \right]
        l > 0, m = 0 \\[2em]
        \qquad m \ne 0
where we define
    \gamma \equiv 1 - \exp\left[ \mu + \frac{1}{2}\nu\right]

\subsection{Second moment}
%
The second moment of the distribution of $y_{lm}$ over $r$ and $b$ is
%
\begin{align}
    \label{eq:Erb2}
    \begin{split}
        \mathrm{E} \big[ y_{lm} y_{l'm'} \big] &=
        \int\limits_0^\infty \int\limits_0^1
        y_{lm} (r, 1 - b)
        y_{l'm'} (r, 1 - b)
        p \big(r, 1 - b \, \big| \, \alpha, \beta, \mu, \nu \big)
        \mathrm{d} r
        \,
        \mathrm{d} b
        \\
        &=
        \int\limits_0^\infty
        \mathrm{E}_r \big[ y_{lm} y_{l'm'} \big]
        p \big(b \, \big| \, \mu, \nu \big)
        \mathrm{d} b
        \quad,
    \end{split}
    %
    \\
    \intertext{where}
    \label{eq:Er2}
    %
    \mathrm{E}_r \big[ y_{lm} y_{l'm'} \big]
     & =
    \int\limits_0^1
    y_{lm} (r, 1 - b)
    y_{l'm'} (r, 1 - b)
    p \big(r \, \big| \, \alpha, \beta \big)
    \mathrm{d} r
    \quad.
\end{align}
%
Expanding Equation~(\ref{eq:Er2}), we obtain
%
\begin{align}
    \mathrm{E}_r \big[ y_{lm} y_{l'm'} \big]
     & =
    \begin{dcases}
        \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
        \int\limits_0^1
        \left(
        1 - \dfrac{(1 - b) c r}{2 (1 + c r)}
        \right)^2
        r^{\alpha - 1}
        (1 - r)^{\beta - 1}
        \mathrm{d} r
         &
        \quad\quad
        \parbox[t]{\linewidth}{$l = l',              \\m = m' = 0$}
        \\[2em]
        %
        -\dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
        \int\limits_0^1
        \left(1 - \dfrac{(1 - b) c r}{2 (1 + c r)}\right)
        \\
        \qquad\qquad\quad\quad\times
        \dfrac{(1 - b) c r \left( 2 + c r \right) r^{\alpha - 1}
            (1 - r)^{\beta - 1}}
        {2 \sqrt{2l + 1} (1 + c r)^{l + 1}}
        \mathrm{d} r
         &
        \quad\quad
        \parbox[t]{\linewidth}{\vspace{-3em}$l > 0,  \\l' = 0, \\m = m' = 0$}
        \\[2em]
        %
        -\dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
        \int\limits_0^1
        \left(1 - \dfrac{(1 - b) c r}{2 (1 + c r)}\right)
        \\
        \qquad\qquad\quad\quad\times
        \dfrac{(1 - b) c r \left( 2 + c r \right) r^{\alpha - 1}
            (1 - r)^{\beta - 1}}
        {2 \sqrt{2l' + 1} (1 + c r)^{l' + 1}}
        \mathrm{d} r
         &
        \quad\quad
        \parbox[t]{\linewidth}{\vspace{-3em}$l = 0,  \\l' > 0, \\m = m' = 0$}
        \\[2em]
        %
        -\dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
        \int\limits_0^1
        \dfrac{(1 - b)^2 c^2 r^2 \left( 2 + c r \right)^2 r^{\alpha - 1}
            (1 - r)^{\beta - 1}}
        {4 \sqrt{(2l + 1)(2l' + 1)} (1 + c r)^{l + l' + 2}}
        \mathrm{d} r
         &
        \quad\quad
        \parbox[t]{\linewidth}{\vspace{-1.5em}$l > 0 \\l' > 0 \\m = m' = 0$}
        \\[2em]
        %
        0
         &
        \quad\quad m, m' \ne 0
        \quad.
    \end{dcases}
\end{align}

>>>>

\subsection{Distributions over $r$ and $\delta$}
%
We model the normalized spot radius $r$ as a random variable drawn from
a Beta distribution, whose PDF is given by
%
\begin{align}
    p \big(r \, \big| \, \alpha, \beta \big)
     & =
    \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
    r^{\alpha - 1}
    (1 - r)^{\beta - 1}
    \quad,
\end{align}
%
where $\Gamma$ is the Gamma function and $\alpha$ and $\beta$ are
hyperparameters characterizing the shape of the distribution.
%
Since we require the brghtness $b = 1 - \delta$ at the center of the spot to
be non-negative, it is convenient to parametrize the distribution of $b$
instead of $\delta$. We model $b$ as a random variable drawn from a log-normal
distribution:
%
\begin{align}
    \label{eq:lognormal}
    p \big(b \, \big| \, \mu, \nu \big)
     & =
    \frac{1}{b\sqrt{2\pi\nu}}
    \exp\left[
        -\dfrac{\left(\ln b - \mu\right)^2}{2\nu}
        \right]
    \quad,
\end{align}
%
where $\mu$ and $\nu$ are the mean and variance of the log-normal, respectively.
The joint distribution over $r$ and $b$ is therefore separable,
by construction:
%
\begin{align}
    p \big( r, b \, \big| \, \alpha, \beta, \mu, \nu \big)
     & =
    p \big(r \, \big| \, \alpha, \beta \big)
    \,
    p \big(b \, \big| \, \mu, \nu \big)
    \quad.
\end{align}
%

\subsection{First moment}
%
The first moment of the distribution of $y_{lm}$ over $r$ and $b$ is
%
\begin{align}
    \label{eq:Erb}
    \begin{split}
        \mathrm{E} \big[ y_{lm} \big] &=
        \int\limits_0^\infty \int\limits_0^1
        y_{lm} (r, 1 - b)
        p \big(r, 1 - b \, \big| \, \alpha, \beta, \mu, \nu \big)
        \mathrm{d} r
        \,
        \mathrm{d} b
        \\
        &=
        \int\limits_0^\infty
        \mathrm{E}_r \big[ y_{lm} \big]
        p \big(b \, \big| \, \mu, \nu \big)
        \mathrm{d} b
        \quad,
    \end{split}
    %
    \\
    \intertext{where}
    \label{eq:Er}
    %
    \mathrm{E}_r \big[ y_{lm} \big]
     & =
    \int\limits_0^1
    y_{lm} (r, 1 - b)
    p \big(r \, \big| \, \alpha, \beta \big)
    \mathrm{d} r
    \quad.
\end{align}
%
Expanding Equation~(\ref{eq:Er}), we obtain
%
\begin{align}
    \mathrm{E}_r \big[ y_{lm} \big]
     & =
    \begin{dcases}
        \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
        \int\limits_0^1
        \left(
        1 - \dfrac{(1 - b) c r}{2 (1 + c r)}
        \right)
        r^{\alpha - 1}
        (1 - r)^{\beta - 1}
        \mathrm{d} r
         &
        \qquad
        l = m = 0    \\[2em]
        -\dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
        \int\limits_0^1
        \dfrac{(1 - b) c r \left( 2 + c r \right)r^{\alpha - 1}
        (1 - r)^{\beta - 1}}
        {2 \sqrt{2l + 1} (1 + c r)^{l + 1}}
        \mathrm{d} r
         &
        \qquad
        l > 0, m = 0 \\[2em]
        0
         &
        \qquad m \ne 0
        \quad.
    \end{dcases}
\end{align}
%
Thanks to the dependence of the spherical harmonic coefficients on only
powers of $r$ and $(1 + cr)$, the integrals above may be expressed in closed
form in terms of the hypergeometric function ${_2F_1}$:
%
\begin{align}
    \mathrm{E}_r \big[ y_{lm} \big]
     & =
    \begin{dcases}
        1 -
        \dfrac{(1 - b) \alpha c}{2(\alpha + \beta)} G_0^1
         &
        \qquad
        l = m = 0    \\[2em]
        -\dfrac{(1 - b) \alpha c}{\sqrt{2l + 1} (\alpha + \beta)}
        \left[
            G_l^1 + \dfrac{(\alpha + 1)c}{2(\alpha + \beta + 1)} G_l^2
            \right]
         &
        \qquad
        l > 0, m = 0 \\[2em]
        0
         &
        \qquad m \ne 0
    \end{dcases}
\end{align}
%
where we define
%
\begin{align}
    G_j^k & \equiv {_2F_1}(j + 1, \alpha + k; \alpha + \beta + k; -c)
    \quad.
\end{align}
%
Note that for a given value of $\alpha$, $\beta$, and $c$, we need only compute
$G_0^1$, $G_1^1$, $G_0^2$, and $G_1^2$ directly, since the remaining terms may be
obtained recursively:
%
\begin{align}
    G_l^k & =
    \dfrac{\alpha + \beta + k - l}{2l}G_{l - 2}^k
    - \dfrac{2\alpha + \beta + 2k - 3l}{2l}G_{l - 1}^k
    \quad.
\end{align}
%
We may now evaluate the integral in Equation~(\ref{eq:Erb}), with
PDF given by Equation~(\ref{eq:lognormal}).
Most of the terms in the integrands are constants, so the
integrals are straightforward to evaluate and may also
be expressed in closed form:
%
\begin{align}
    \mathrm{E} \big[ y_{lm} \big]
     & =
    \begin{dcases}
        1 -
        \dfrac{\gamma_1 \alpha c}{2(\alpha + \beta)} G_0^1
         &
        \qquad
        l = m = 0    \\[2em]
        -\dfrac{\gamma_1 \alpha c}{\sqrt{2l + 1} (\alpha + \beta)}
        \left[
            G_l^1 + \dfrac{(\alpha + 1)c}{2(\alpha + \beta + 1)} G_l^2
            \right]
         &
        \qquad
        l > 0, m = 0 \\[2em]
        0
         &
        \qquad m \ne 0
    \end{dcases}
\end{align}
%
where we define
%
\begin{align}
    \gamma_1 \equiv 1 - \exp\left[ \mu + \frac{1}{2}\nu\right]
    \quad.
\end{align}
%

\subsection{Second moment}

@@

The second moment of the distribution of $y_{lm}$ over $r$ and $b$ is

@@

    \label{eq:Erb2}

@@

        \mathrm{E} \big[ y_{lm} y_{l'm'} \big] &=

@@

        y_{l'm'} (r, 1 - b)

@@

        \mathrm{E}_r \big[ y_{lm} y_{l'm'} \big]

@@

    \label{eq:Er2}

@@

Expanding Equation~(\ref{eq:Er2}), we obtain

@@

        \right)^2

@@

        \quad\quad
        \parbox[t]{\linewidth}{$l = l',               \\m = m' = 0$}
        \\[2em]

@@

        \left(1 - \dfrac{(1 - b) c r}{2 (1 + c r)}\right)
        \qquad\qquad\quad\quad\times
        \dfrac{(1 - b) c r \left( 2 + c r \right) r^{\alpha - 1}

@@

        \quad\quad
        \parbox[t]{\linewidth}{\vspace{-3em}$l > 0,   \\l' = 0, \\m = m' = 0$}
        \\[2em]
        \left(1 - \dfrac{(1 - b) c r}{2 (1 + c r)}\right)
        \qquad\qquad\quad\quad\times
        \dfrac{(1 - b) c r \left( 2 + c r \right) r^{\alpha - 1}
        {2 \sqrt{2l' + 1} (1 + c r)^{l' + 1}}
        \quad\quad
        \parbox[t]{\linewidth}{\vspace{-3em}$l = 0,   \\l' > 0, \\m = m' = 0$}
        \\[2em]
        \dfrac{(1 - b)^2 c^2 r^2 \left( 2 + c r \right)^2 r^{\alpha - 1}
        {4 \sqrt{(2l + 1)(2l' + 1)} (1 + c r)^{l + l' + 2}}
        \quad\quad
        \parbox[t]{\linewidth}{\vspace{-1.5em}$l > 0, \\l' > 0, \\m = m' = 0$}
        \\[2em]

@@

        \quad\quad m, m' \ne 0

@@

These integrals again reduce to closed form expressions:

@@

        1 + \frac{\left(\alpha +1\right) \alpha
            \left(1-b\right)^2 c^2}{4 \left(c+1\right)
            \left(\alpha + \beta \right)}
        \\[0.75em]
        \quad\quad\quad\quad\quad
        -\frac{\alpha  \left(1-b\right) c}{\alpha +\beta }
        \\[0.75em]
        \quad\quad\quad\quad\quad
        \times
        \bigg[G_0^1
            +
            \frac{\left(\alpha +1\right) \left(1-b\right)
                c \left(\alpha +\beta +\alpha  c+c\right)}{4 \left(c+1\right)
                \left(\alpha +\beta +1\right)} G_0^2\bigg]

@@

        \quad\quad
        \parbox[t]{\linewidth}{\vspace{-4.75em}$l = l', \\m = m' = 0$}
        \\[2em]
        -\frac{\alpha (1-b) c}{2 \sqrt{2 l+1} (\alpha +\beta )}
        \\[0.75em]
        \quad\quad\quad\quad\quad
        \times
        \bigg[
            2G_l^1
            + \dfrac{c(\alpha + 1)}{\alpha + \beta + 1} G_l^2
            \\[0.75em]

@@

        -\dfrac{c(1-b)(\alpha + 1)}{\alpha + \beta + 1} G_{l+1}^2
        \\[0.75em]
        -\dfrac{c^2(1-b)(\alpha + 1)(\alpha + 2)}{2(\alpha + \beta + 1)(\alpha + \beta + 2)} G_{l+1}^3
        \bigg]

@@

        \quad\quad
        \parbox[t]{\linewidth}{\vspace{-7.5em}$l > 0,   \\l' = 0, \\m = m' = 0$}
        \\[2em]
        \textrm{same as previous case, but with $l \rightarrow l'$}
        \quad\quad
        \parbox[t]{\linewidth}{\vspace{-2em}$l = 0,     \\l' > 0, \\m = m' = 0$}
        \\[2em]
        \frac{\alpha  (\alpha +1) (1-b)^2 c^2}{(\alpha +\beta )
            (\alpha +\beta +1) \sqrt{(2 l+1) (2 p+1)}}
        \\[0.75em]
        \quad\quad\quad\quad\quad
        \times
        \bigg[
        G_{l+l'+1}^2
        +
        \frac{c(\alpha + 2)}{\alpha + \beta + 2}G_{l+l'+1}^3
        \\[0.75em]

@@

        +
        \frac{c^2(\alpha + 2)(\alpha + 3)}{4(\alpha + \beta + 2)(\alpha + \beta + 3)}G_{l+l'+1}^4
        \bigg]
        \quad\quad
        \parbox[t]{\linewidth}{\vspace{-5.25em}$l > 0,  \\l' > 0, \\m = m' = 0$}
        \\[2em]

@@

The $G_j^k$ functions may again be computed recursively, although we now also
require the direct evaluation of $G_0^3$, $G_1^3$, $G_0^4$, and $G_1^4$ as
our lower boundary conditions.

The final step is to evaluate the integral in Equation~(\ref{eq:Er2}), given
the PDF in Equation~(\ref{eq:lognormal}):
%
\begin{align}
    \mathrm{E} \big[ y_{lm} y_{l'm'} \big]
     & =
    \begin{dcases}
        1 + \frac{\left(\alpha +1\right) \alpha
            \gamma_2 c^2}{4 \left(c+1\right)
            \left(\alpha + \beta \right)}
        \\[0.75em]
        \quad\quad\quad\quad\quad
        -\frac{\alpha  \gamma_1 c}{\alpha +\beta }
        \\[0.75em]
        \quad\quad\quad\quad\quad
        \times
        \bigg[G_0^1
            +
            \frac{\left(\alpha +1\right) \gamma_2
                c \left(\alpha +\beta +\alpha  c+c\right)}{4 \gamma_1 \left(c+1\right)
                \left(\alpha +\beta +1\right)} G_0^2\bigg]
         &
        \quad\quad
        \parbox[t]{\linewidth}{\vspace{-4.75em}$l = l', \\m = m' = 0$}
        \\[2em]
        %
        -\frac{\alpha \gamma_1 c}{2 \sqrt{2 l+1} (\alpha +\beta )}
        \\[0.75em]
        \quad\quad\quad\quad\quad
        \times
        \bigg[
            2G_l^1
            + \dfrac{c(\alpha + 1)}{\alpha + \beta + 1} G_l^2
            \\[0.75em]
        \qquad
        -\dfrac{c\gamma_2(\alpha + 1)}{\gamma_1(\alpha + \beta + 1)} G_{l+1}^2
        \\[0.75em]
        \qquad
        -\dfrac{c^2\gamma_2(\alpha + 1)(\alpha + 2)}{2\gamma_1(\alpha + \beta + 1)(\alpha + \beta + 2)} G_{l+1}^3
        \bigg]
         &
        \quad\quad
        \parbox[t]{\linewidth}{\vspace{-7.5em}$l > 0,   \\l' = 0, \\m = m' = 0$}
        \\[2em]
        %
        \textrm{same as previous case, but with $l \rightarrow l'$}
         &
        \quad\quad
        \parbox[t]{\linewidth}{\vspace{-2em}$l = 0,     \\l' > 0, \\m = m' = 0$}
        \\[2em]
        %
        \frac{\alpha  (\alpha +1) \gamma_2 c^2}{(\alpha +\beta )
            (\alpha +\beta +1) \sqrt{(2 l+1) (2 p+1)}}
        \\[0.75em]
        \quad\quad\quad\quad\quad
        \times
        \bigg[
        G_{l+l'+1}^2
        +
        \frac{c(\alpha + 2)}{\alpha + \beta + 2}G_{l+l'+1}^3
        \\[0.75em]
        \qquad
        +
        \frac{c^2(\alpha + 2)(\alpha + 3)}{4(\alpha + \beta + 2)(\alpha + \beta + 3)}G_{l+l'+1}^4
        \bigg]
         &
        \quad\quad
        \parbox[t]{\linewidth}{\vspace{-5.25em}$l > 0,  \\l' > 0, \\m = m' = 0$}
        \\[2em]
        %
        0
         &
        \quad\quad m, m' \ne 0
        \quad,
    \end{dcases}
\end{align}
%
where we define
%
\begin{align}
    \gamma_2 \equiv 1
    - 2\exp\bigg[ \mu + \frac{1}{2}\nu\bigg]
    + \exp\bigg[ 2\mu + 2\nu\bigg]
    \quad.
\end{align}
